Memory-Efficient Mixed-Precision Implementations\\for Robust Explicit Model Predictive Control

### 2019-04-12 emsoft-2019/Error_Analysis.tex
Various changes throughout.

In this section, we present our error analysis assuming that a fixed-point word lengths is given.
Then in the next section, we explain our optimization algorithm which determines suitable word
lengths fully automatically.% in~\autoref{sec:Controller_Synthesis}.
While we do not employ the error analysis independently of the controller and
precision synthesis, we note that it is an alternative way of bounding errors to
existing methods~\cite{imperialrmpc}, which phrase the roundoff error computation as a (relaxed)
mixed integer programming problem.
%\Sadegh{Alternative in what sense? What is the major difference between our analysis and \cite{imperialrmpc}?}

For simplicity of notation, we assume in this section a uniform word length $p$,
but our algorithm (and implementation) is applied to mixed-precision to reduce the memory usage.

@@

We have to ensure that the sum of these two errors remain below the bound $\Delta$
used in the design of the robust explicit MPC:
\begin{align}
  \max\{\|v_{i}-v_{j}\|,\,&\forall i,j,\,\, i\neq j, \mathcal{R}_{i}\cap\mathcal{R}_{j}\neq\emptyset\} \nonumber\\
  &+  \max\{err(v_{i})_{p},\, i=1,2,\ldots,P\} \le \Delta.
  \label{eq:delta}
\end{align}
%where $i,j$ are the indexes of two distinct regions $\mathcal{R}_{i}$ and $\mathcal{R}_{j}$, 
The first term is the error of incorrect region selection. The maximum is taken over all neighboring regions $\mathcal R_i,\mathcal R_j$ recognisable by having non-empty intersection. The function norm $\|v_{i}-v_{j}\|$ is taken by maximizing over all possible values of the state $x_k$. 
$err(v_{i})_{p}$ in
the second term captures the approximation error in the computation of the affine function $v_{i}$ when a fixed precision $p$ is used.
%$\mathit{neighbor}(\mathcal{R}_{i},\mathcal{R}_{j})$
%states that two regions share at least a point, $v_{i}$ is the control action
%for controller $i$, and $err(v_{i})_{p}$ captures the approximation error for
%computing the control action $v_{i}$ in fixed precision $p$.

@@

While $\varepsilon_{A/D}$ is intrinsic to the capabilities of the device,
$\varepsilon_{Q}$ depends on the precision used to store the boundaries:
\begin{equation}\label{eq:quantizationlines}
  \varepsilon_{Q} = \|(H_i - \hat{H}_i) \hat{x} + (K_i - \hat{K}_i)\|
\end{equation}
where $\hat{x}$ is the finite-precision value of $x$.
%\eva{For which index is this evaluated, $i$, $j$ or both?}
That is, $\varepsilon_{Q}$ bounds the distance between any hyperplane in infinite
precision (H and K), and its counter-part quantized in $p$ bits precision
($\hat{H}_{p}$ and $\hat{K}_{p}$).
This second error can be tuned providing a trade-off between accuracy and
memory storage required.

@@

Due to the affinity of function
$v_{i}-v_{j}$, and because we defined the tube as a convex region surrounding the corresponding hyperplane, it suffices
to evaluate the function only at the corner points of the tube, and keep the
result with the maximal magnitude. 
To evaluate the incorrect region selection error across
the corner points, we need to first locate the corner points as well as all the
regions which share those corner points. 
Each region $R_i$ is represented by a set of contraints. We first extract the
set of vertices of $R_i$ using the open source Matlab library
$\mathit{lcon2vert}$~\cite{??}.
For each computed vertex $v_i$, we compute a $n-D$
hypercube with the edge length of $2 \varepsilon$. 
Each of the $2^n$ vertices of this hypercube is counted as a corner point, on which
we evaluate $\|u_i - u_j\|$.

@@

$\varepsilon_Q$ and $err(v_{i})_{p}$ are computed by Daisy. For this we encode
the expressions in~\autoref{eq:quantizationlines} and \autoref{eq:fperror}
respectively as straight-line arithmetic expressions and specify the constraints
on the domain of $\statevar$ in the precondition.
Daisy performs a dataflow analysis to determine the finite-precision roundoff
errors. 
% This error computation differs from the one employed by
% \citet{imperialrmpc} which phrases the roundoff error computation as a (relaxed)
% mixed integer programming problem.

>>>>

In this section, we present our error analysis assuming that a fixed word length $p$ is given.
Then in the next section, we explain our optimization algorithm which determines suitable word
lengths for different variables fully automatically. % in~\autoref{sec:Controller_Synthesis}.
%
% For simplicity of notation, we assume in this section a uniform word length $p$,
% but our algorithm (and implementation) is applied to mixed-precision to reduce the memory usage.

@@

We have to ensure that the sum of these two errors remain below the bound $\Delta$
used in the design of the robust explicit MPC:
\begin{align}
  \max\{\|v_{i}-v_{j}\|,\,&\forall i,j,\,\, i\neq j, \mathcal{R}_{i}\cap\mathcal{R}_{j}\neq\emptyset\} \nonumber\\
  &+  \max\{err(v_{i})_{p},\, i=1,2,\ldots,P\} \le \Delta.
  \label{eq:delta}
\end{align}
%where $i,j$ are the indexes of two distinct regions $\mathcal{R}_{i}$ and $\mathcal{R}_{j}$, 
The first term is the error of incorrect region selection. The maximum is taken over all neighboring regions 
$\mathcal R_i,\mathcal R_j$ (two regions are neighboring if their intersection is non-empty).
The function norm $\|v_{i}-v_{j}\|$ is taken by maximizing over all possible values of the state $x_k$. 
$err(v_{i})_{p}$ in
the second term captures the approximation error in the computation of the affine function $v_{i}$ when a fixed precision $p$ is used.
%$\mathit{neighbor}(\mathcal{R}_{i},\mathcal{R}_{j})$
%states that two regions share at least a point, $v_{i}$ is the control action
%for controller $i$, and $err(v_{i})_{p}$ captures the approximation error for
%computing the control action $v_{i}$ in fixed precision $p$.

@@

While $\varepsilon_{A/D}$ is intrinsic to the capabilities of the device,
$\varepsilon_{Q}$ depends on the precision used to store the boundaries:
\begin{equation}\label{eq:quantizationlines}
  \varepsilon_{Q} = \|(H_i - \hat{H}_i) \hat{x} + (K_i - \hat{K}_i)\|
\end{equation}
where $\hat{x}$ is the finite-precision value of $x$.
%\eva{For which index is this evaluated, $i$, $j$ or both?}
That is, $\varepsilon_{Q}$ bounds the distance between any hyperplane in infinite
precision ($H$ and $K$ matrices), and its counter-part quantized by $p$ bits 
($\hat{H}_{p}$ and $\hat{K}_{p}$).
This second error can be tuned providing a trade-off between accuracy and
memory storage required.

@@

Since $v_{i}-v_{j}$ is also affine, and because we defined the tube as a convex region 
surrounding the corresponding hyperplane, it suffices
to evaluate the function only at the corner points of the tube, and keep the
result with the maximal magnitude. 
To evaluate the incorrect region selection error across
the corner points, we need to first locate the corner points as well as all the
regions which share those corner points. 
Each region $R_i$ is represented by a set of contraints. We first extract the
set of vertices of $R_i$ using the open source Matlab library
$\mathit{lcon2vert}$~\cite{lcon2vertMatlab}.
For each computed vertex $v_i$, we compute a $n$-dimensional
hypercube with the edge length of $2 \varepsilon$. 
Each of the $2^n$ vertices of this hypercube is counted as a corner point, on which
we evaluate $\|u_i - u_j\|$.

@@

The error terms $\varepsilon_Q$ and $err(v_{i})_{p}$ are computed by Daisy. For this we encode
the expressions in~\autoref{eq:quantizationlines} and \autoref{eq:fperror}
respectively as straight-line arithmetic expressions and specify the constraints
on the domain of $\statevar$ in the precondition.
Daisy performs a dataflow analysis to determine the finite-precision roundoff
errors. 
% This error computation differs from the one employed by
% \citet{imperialrmpc} which phrases the roundoff error computation as a (relaxed)
% mixed integer programming problem.

### 2019-04-12 emsoft-2019/Motiv.tex
Various changes throughout.

% !TEX root = main.tex
\section{Overview}
\label{sec:example}
\begin{figure}
\centering
	\tikzstyle{block} = [draw, rectangle, minimum height=4em, minimum width=6em]
	\tikzstyle{block1} = [draw, rectangle, minimum height=2em, minimum width=6em]
	\tikzstyle{block2} = [draw, rectangle, minimum height=1em, minimum width=4em]
	\tikzstyle{block3} = [draw, rectangle, minimum height=3.2em, minimum width=4em]
	\tikzstyle{input} = [coordinate]
	\tikzstyle{output} = [coordinate]
	\tikzstyle{pinstyle} = [pin edge={to-,thin,black}]
	{

@@

% Model Predictive Control (MPC) refers to the class of control strategies that design control actions by solving a finite horizon optimal control problem at each sampling instant. While each optimization gives a sequence of optimal control actions, only the first action is applied to the process. The same procedure is applied in the next time instance after receiving the updated value of the process state. MPC has shown to be successful in industrial applications since it is able to handle large number of states and control actions and automatically considers constraints in the optimization.
% 
% The optimization of each time instance uses a dynamic model of the process to predict the future states and find the optimal sequence of actions that satisfies all the constraints. The objective function of the optimization is usually linear or quadratic. The resulting optimization problem can be cast as a linear program (LP) or quadratic program (QP), respectively, if the dynamic model is linear. For hybrid models with linear dynamics in each mode, the optimization problem can be cast as a mixed integer linear or quadratic program (MILP/MIQP).

\begin{figure}[h]
	\begin{tabular}{cc}
	\includegraphics[width=2cm,height=3cm]{Figs/inverted_pendulum.pdf}&	\includegraphics[width=6cm,height=4cm]{Figs/regs.jpg}\\
	(a)&(b)
	\end{tabular}
	\caption{(a) Inverted pendulum; (b) 2-D plot of polyhedral partitions for an explicit MPC applied to the inverted pendulum.}
	\label{fig:inverted_pendulum}
\end{figure}

@@

For our example, we choose $\Delta_0=0.01$ and $\mathcal W = \{0\}$. 
We use Matlab to find a robust explicit MPC with disturbance set $\mathcal W_{\Delta_0}$. 
% \RM{the following setence is not understandable. This is the first time we introduce incorrect region selection.
% Who is Daisy?}
% \RM{A better version is to informally define incorrect region selection. Say that we run an algorithm for incorrect region
% selection to bound the possible error. 
% Break the sentence into two. Introduce Daisy as a mixed-precision tuning tool.}

The output of robust EMPC given by Matlab decomposes the control domain into a
finite number of polyhedral domains together with an affine map for each domain. At each
time step, the state vector $x_k$ is read (or estimated) and based on the domain
that it belongs to, the corresponding control output is computed. 
However, the polyhedral regions are implemented in finite precision so that
it is possible that due to quantization errors the domain is selected incorrectly.
We refer to the resulting error as incorrect region selection error.
% However, due to error by fixed-point implementation, it is possible that the domain is selected
% incorrectly. This causes an error term which we refer to it as incorrect region
% selection error.

In our example, we compute an incorrect region selection error $e = 0.019$, which is larger
than $\Delta_0=0.01$. We therefore increase the disturbance bound by a fixed amount to
$\Delta_1 = 0.05$ and enlarge the disturbance set as $\mathcal W_{\Delta_1}$. 
%
% Over the high-precision controller, we compute the error due to an incorrect region selection $e$. In case that $e<\Delta_0$, we would like to find the smallest number of bits required for implementing the controller such that we are still within the safe bound $\Delta_0$. To do so, we ask Daisy, which is a mixed-precision tuning tool, to provide us with a mixed-precision scheme such that the associated implementation error is at most $\Delta_0$, that is,
% at all points in the control domain, the difference between the infinite-precision controller function
% and its mixed-precision implementation is at most $\Delta_0$.
%
%Unfortunately, $\Delta_0=0.01$ is not realizable, as the region selection error is $0.019$ and hence $e>\Delta_0$. 
%always greater than zero either due to selecting the wrong region in the affine controller function or due to the finite-precision computation of the control action.
%\eva{Is this the error due to mixed-precision or due to wrong regions?}
% Therefore, we define $\Delta_1 := \Delta_0 + \varepsilon_{SAFE}$ with $\varepsilon_{SAFE}$ being a fixed threshold (we chose $\varepsilon_{SAFE} = 0.04$) and enlarge the disturbance set as $\mathcal W_{\Delta_1}$. 
Solving the robust explicit MPC problem for this enlarged disturbance set gives the new controller
with 14 regions, shown in~\autoref{fig:inverted_pendulum}(b).
A sample region with the corresponding affine map is shown below:
%Following demostrates one sample mapping, including the activation condition and its corresponding affine function
\begin{equation*}
\resizebox{.9\hsize}{!}{$
\begin{bmatrix}
-0.005  &0.999\\
-0.998  &-0.049\\
0.005  &-0.999\\
0.998   &0.049
\end{bmatrix}x_k\leq 
\begin{bmatrix}
0.416\\
3.157\\
0.617\\
-0.0124
\end{bmatrix}
\Rightarrow
u(x_k)=\begin{bmatrix}
0.05&-9.67
\end{bmatrix}x_k+4.02$}
\end{equation*}
where the output of robust EMPC at state $x_k$ is denoted by $u(x_k)$. 
The remaining mappings have a similar structure.
For this new controller, the incorrect region selection error is still
approximately $e = 0.019$, which is now below the disturbance bound $\Delta_1$,
so that we can continue with the precision assignment.

@@

Next, we use the precision tuning tool Daisy to provide us with a mixed-precision 
implementation for all the parameters which satisfies the error bound $0.031$,
which is given by the remaining error budget $\Delta_1 - e = 0.05 - 0.019$.  
The mixed-precision implementation returned by Daisy requires $6084$ bits in total 
for the corresponding robust explicit MPC.
% Since this error bound is less than our robustness margin $\Delta_1 = 0.05$, we terminate the loop
% and return this implementation.
The mixed-precision implementation takes about $10\%$ less memory compared 
to the smallest uniform precision implementation that respects $\Delta_1$,
and about $57\%$ less memory than an implementation that uniformly uses
32 bits.

>>>>

% !TEX root = main.tex
\begin{figure}[t]
\centering
	\tikzstyle{block} = [draw, rectangle, minimum height=4em, minimum width=6em]
	\tikzstyle{block1} = [draw, rectangle, minimum height=2em, minimum width=6em]
	\tikzstyle{block2} = [draw, rectangle, minimum height=1em, minimum width=4em]
	\tikzstyle{block3} = [draw, rectangle, minimum height=3.2em, minimum width=4em]
	\tikzstyle{input} = [coordinate]
	\tikzstyle{output} = [coordinate]
	\tikzstyle{pinstyle} = [pin edge={to-,thin,black}]
	{

@@

\section{Overview}
\label{sec:example}

\begin{figure}[t]
	\begin{tabular}{cc}
	\includegraphics[width=2cm,height=3cm]{Figs/inverted_pendulum.pdf}&	\includegraphics[width=6cm,height=4cm]{Figs/regs.jpg}\\
	(a)&(b)
	\end{tabular}
	\caption{(a) Inverted pendulum; (b) 2-D plot of polyhedral partitions for an explicit MPC applied to the inverted pendulum.}
	\label{fig:inverted_pendulum}
\end{figure}

@@

For our example, we choose $\Delta_0=0.01$ and $\mathcal W = \{0\}$. 
We use Matlab to find a robust explicit MPC with disturbance set $\mathcal W_{\Delta_0}$. 
%
The output of robust EMPC given by Matlab decomposes the control domain into a
finite number of polyhedral domains together with an affine map for each domain. 
At each time step, the state vector $x_k$ is read (or estimated) and based on the domain
that it belongs to, the corresponding control output is computed. 
However, the polyhedral regions are implemented in finite precision so that
it is possible that due to quantization errors the domain is selected incorrectly, and therefore
a different affine map is computed for the control.
We refer to the resulting error as \emph{incorrect region selection} error.
% However, due to error by fixed-point implementation, it is possible that the domain is selected
% incorrectly. This causes an error term which we refer to it as incorrect region
% selection error.
In our example, we compute an incorrect region selection error $e = 0.019$, which is larger
than $\Delta_0=0.01$. We therefore increase the disturbance bound by a fixed amount to
$\Delta_1 = 0.05$ and enlarge the disturbance set as $\mathcal W_{\Delta_1}$.

%
Solving the robust explicit MPC problem for this enlarged disturbance set gives the new controller
with 14 regions, shown in~\autoref{fig:inverted_pendulum}(b).
A sample region with the corresponding affine map is shown below:
\begin{equation*}
\resizebox{.9\hsize}{!}{$
\begin{bmatrix}
-0.005  &0.999\\
-0.998  &-0.049\\
0.005  &-0.999\\
0.998   &0.049
\end{bmatrix}x_k\leq 
\begin{bmatrix}
0.416\\
3.157\\
0.617\\
-0.0124
\end{bmatrix}
\Rightarrow
u(x_k)=\begin{bmatrix}
0.05&-9.67
\end{bmatrix}x_k+4.02$}
\end{equation*}
where the output of robust EMPC at state $x_k$ is denoted by $u(x_k)$. 
The remaining mappings have a similar structure.
For this new controller, the incorrect region selection error is 
approximately $e = 0.019$, which is now below the disturbance bound $\Delta_1$,
so that we can continue with the precision assignment.

@@

Next, we use the precision tuning tool Daisy \cite{Daisy} to provide a mixed-precision 
implementation for all the parameters which satisfies the error bound $\Delta_1  - e = 0.05 - 0.019 = 0.031$.
The mixed-precision implementation returned by Daisy requires $6084$ bits.
Thus, the mixed-precision implementation takes about $10\%$ less memory compared 
to the smallest uniform precision implementation that respects $\Delta_1$,
and about $57\%$ less memory than an implementation that uniformly uses
32 bits.

### 2019-04-12 emsoft-2019/intro.tex
Various changes throughout.

One way to tackle these problems is through \emph{explicit MPC} \cite{Bemporad:2002,Alessio2009}, which formulates the optimization
problem but computes offline a symbolic representation of the solution as a function of the state.
At run-time, the solution is evaluated on the current state as in conventional control.
For example, for linear time invariant models with linear constraints and quadratic costs, the optimization
problem for MPC can be modeled as a quadratic program, and explicit MPC techniques solve the optimization problem
using multi-parametric programming techniques.
The explicit solution is representable as a partition of the controller domain into a number of polyhedral regions,
and an affine map for each region.
The implementation of explicit MPC stores a lookup table of the affine maps and performs a number of affine computations
online to find the appropriate region for the current state and to evaluate the control.  
Explicit MPC thus expands the class of systems being controlled by MPC strategies, by taking out the need to compute fast online
or to certify a complex optimization routine.

However, there are still bottlenecks in implementing explicit MPC on resource-constrained embedded micro-controllers.
First, implementations of explicit MPCs can suffer from large memory usage, because the solution of the optimization
problem can involve many (often hundreds) of regions.
Since the memory consumption grows linearly with the number of regions, 
this can be a limiting factor in using explicit MPC on resource-constrained micro-controllers.
Second, many low-end micro-controllers only support fixed-point arithmetic.
Errors in the controller implementation are inversely proportional to the number of bits used for representing each variable~\cite{Anta10}. 
An implementation of explicit MPC has to be \emph{robust} to 
implementation errors to be able to enforce hard constraints on the states at run time~\cite{imperialrmpc}.

In this paper, we consider the problem of implementing explicit MPC on low-end microcontrollers with fixed-point arithmetic
in a \emph{memory-efficient} and \emph{robust} way.
We propose an automatic controller and mixed-precision implementation co-design
technique that computes mixed-precision assignments which minimize bitwidths for all
variables while ensuring the resulting implementation error remains within the
robustness margin.

Our proposed method iteratively solves a robust version of MPC 
which explicitly considers finite-precision implementation errors as disturbances
in the dynamics \cite{RamirezC06}; see~\autoref{fig:overview}.
Initially, we estimate a bound $\Delta = \Delta_0$ on the implementation error and solve a
min-max quadratic program for explicit robust MPC \cite{delaPea:2005,RamirezC06}, where the system
model has an explicit disturbance bounded by $\Delta$.
The solution is represented as a piecewise affine map.
\RM{need to add the step on incorrect region selection.}
We use an automated mixed-precision tuning tool to find an approximate implementation of the piecewise affine
map. 
We statically compute the error of the implementation, taking into account errors
in evaluating the affine maps as well as choosing the wrong affine map due to quantization.
If the error is at most $\Delta$, we are done: the implementation satisfies the robustness margin in the model.
If not, we increase the error bound $\Delta$ and run the loop again to find a new robust controller and find its best mixed-precision
implementation.

@@

In summary, our contributions are:
\begin{itemize}
\item We design a static and scalable error analysis algorithm for piecewise affine functions;
\item We provide an end-to-end tuning algorithm for robust explicit MPC;
\item Through a set of standard control benchmarks, we demonstrate that our tuning algorithm
can achieve up to 20\% savings in memory while maintaining robustness.
\end{itemize}

>>>>

One way to tackle these problems is through \emph{explicit MPC} (EMPC) \cite{Bemporad:2002,Alessio2009}, which formulates the optimization
problem but computes offline a symbolic representation of the solution as a function of the state.
At run-time, the solution is evaluated on the current state as in conventional control.
For example, for linear time invariant models with linear constraints and quadratic costs, the optimization
problem for MPC can be modeled as a quadratic program, and EMPC techniques solve the optimization problem
using multi-parametric programming techniques.
The explicit solution is representable as a partition of the controller domain into a number of polyhedral regions,
and an affine map for each region.
The implementation of EMPC stores a lookup table of the affine maps and performs a number of affine computations
online to find the appropriate region for the current state and to evaluate the control.  
Explicit MPC thus expands the class of systems being controlled by MPC strategies, by taking out the need to compute fast online
or to certify a complex optimization routine.

However, there are still bottlenecks in implementing EMPC on resource-constrained embedded micro-controllers.
First, implementations of EMPCs can suffer from large memory usage, because the solution of the optimization
problem can involve many (often hundreds) of regions~\cite{memoryMPC}.
Since the memory consumption grows linearly with the number of regions, 
this can be a limiting factor in using EMPC on resource-constrained micro-controllers.
Second, many low-end micro-controllers only support fixed-point arithmetic.
Errors in the controller implementation are inversely proportional to the number of bits used for representing each variable~\cite{Anta10}. 
An implementation of EMPC has to be \emph{robust} to 
implementation errors to be able to enforce hard constraints on the states at run time~\cite{imperialrmpc}.

In this paper, we consider the problem of implementing EMPC on low-end microcontrollers with fixed-point arithmetic
in a \emph{memory-efficient} and \emph{robust} way.
We propose an automatic controller and mixed-precision implementation co-design
technique that computes mixed-precision assignments which minimize bitwidths for all
variables while ensuring the resulting implementation error remains within the
robustness margin.

Our proposed method iteratively solves a robust version of MPC 
which considers finite-precision implementation errors as disturbances
in the dynamics \cite{RamirezC06}; see~\autoref{fig:overview}.
Initially, we estimate a bound $\Delta = \Delta_0$ on the implementation error and solve a
min-max quadratic program for explicit robust MPC \cite{delaPea:2005,RamirezC06}, where the system
model has an explicit disturbance bounded by $\Delta$.
The solution is represented as a set of polyhedral regions, and an affine map for each region.

Next, we consider the error in the control input arising out of a fixed-point implementation of the solution.
The error has two sources.
First, the fixed point implementation may pick a different polyhedral region (due to imprecision
in checking membership in a polyhedron).
Second, the fixed-precision implementation of the affine function will have a numerical error due to quantization.
We bound both sources of error statically and automatically,  by finding the maximum possible error due to incorrect region selection
as well as a static bound on the error in computing the affine map.
If the error is at most $\Delta$, we know that the implementation satisfies the robustness margin in the model and
we use an automated mixed-precision tuning tool to find an optimal bitwidth allocation that keeps the implementation below the error bound.
If not, we increase the error bound $\Delta$ and run the loop again to find a new robust controller and find its best mixed-precision
implementation.

@@

In summary, our contributions are:
\begin{itemize}
\item We provide an end-to-end automatic tuning algorithm for robust explicit MPC that ensures
hard constraints on the state are met despite numerical errors;
\item We design a static and scalable error analysis algorithm for piecewise affine functions to bound
errors in controller implementations; and
\item Through a set of standard control benchmarks, we demonstrate that our tuning algorithm
can achieve up to 20\% savings in memory while maintaining robustness.
\end{itemize}

### 2019-04-12 emsoft-2019/related.tex
Various changes throughout.

% finite-precision only
\paragraph{Finite-Precision Optimization}
General-purpose techniques for synthesizing fixed-point implementations of 
arithmetic expressions have been developed in the space of embedded systems,
where resource efficiency is generally important.
Some work has used dynamic analyses for estimating
errors~\cite{Gaffar2004,Mallik2007}, which, however, do not provide guarantees.
Several approaches~\cite{Lee2006,Osborne2007,Kinsman2009,Pang2011} use sound
error analysis techniques and perform mixed-precision optimization~\cite{Lee2006,Pang2011},
similar to the approach implemented in the tool Daisy which we use.

>>>>

% finite-precision only
\paragraph{Finite-Precision Optimization}
General-purpose techniques for synthesizing fixed-point implementations of 
arithmetic expressions have been developed in the space of embedded systems,
where resource efficiency is generally important.
Some work has used dynamic analyses for estimating
errors~\cite{Gaffar2004,Mallik2007}, which, however, do not provide guarantees.
Several approaches~\cite{Lee2006,Osborne2007,Kinsman2009,Pang2011} use sound
error analysis techniques and perform mixed-precision optimization~\cite{Lee2006,Pang2011},
similar to the approach implemented in the tool Daisy used in this paper.
An alternate way to bound errors in fixed point implementations of EMPC 
through mixed integer programming is given by~\citet{imperialrmpc}.
